---
title: "p8105_hw1_yx3033"
Author: Yvonne Xie
Data: 2025-09-20
output: github_document
---

## Problem 1
```{r}
library(moderndive)
data("early_january_weather")
head(early_january_weather)
```
The variables include `r names(early_january_weather)`.
This dataset has `r nrow(early_january_weather)` rows and `r ncol(early_january_weather)` columns.
The mean temperature is `r round(mean(early_january_weather$temp), digits =2)`F.

Important variables are:  
`time_hour`:the time of recording (hourly), it's showing year-month-day and time in EST e.g. "2013-01-01 01:00:00 EST"  
`temp`: temperature in F, ranges from 24.08 to 57.92  
`humid`: relative humidity, worte in percentage, 59.4 means 59.4%  

```{r}
library(ggplot2)
```
```{r}
ggplot(early_january_weather, aes(x=time_hour, y=temp, color=humid))+
geom_point() + 
labs(
    title = "Temperature vs Time in Early January",
    x = "Time (hourly)",
    y = "Temperature (F)",
    color = "Humidity"
  )

ggsave("scatterplot_temperature_vs_time.png")
```
 
The scatterplot shows how temperatures fluctuates during early days of January. The temperatures range between mid-20s to mid-50s F. It shows a cycle of waming the cooling throughout the day. Temperature shows an general increasing trend from the beginning of January to mid-January.


## Problem 2
```{r}
library(tidyverse)
df <- tibble(
  random_samp = rnorm(10),
  is_positive = random_samp>0,
  vec_char = c("a", "b", "c", "a","b", "c","a","b", "c","a"),
  fac_vec = factor(c("male","other","female","female","male","other","male","female","male","female"))
)
df
```

