---
title: "p8105_hw1_yx3033"
Author: Yvonne Xie
Data: 2025-09-20
output: github_document
---

## Problem 1
```{r}
library(moderndive)
data("early_january_weather")
head(early_january_weather)
```
The variables include `r names(early_january_weather)`.
This dataset has `r nrow(early_january_weather)` rows and `r ncol(early_january_weather)` columns.
The mean temperature is `r round(mean(early_january_weather$temp), digits =2)`F.

Important variables are:  
`time_hour`:the time of recording (hourly), it's showing year-month-day and time in EST e.g. "2013-01-01 01:00:00 EST"  
`temp`: temperature in F, ranges from 24.08 to 57.92  
`humid`: relative humidity, worte in percentage, 59.4 means 59.4%  

```{r}
library(ggplot2)
```
```{r}
ggplot(early_january_weather, aes(x=time_hour, y=temp, color=humid))+
geom_point() + 
labs(
    title = "Temperature vs Time in Early January",
    x = "Time (hourly)",
    y = "Temperature (F)",
    color = "Humidity"
  )

ggsave("scatterplot_temperature_vs_time.png")
```
 
The scatterplot shows how temperatures fluctuates during early days of January. The temperatures range between mid-20s to mid-50s F. It shows a cycle of waming the cooling throughout the day. Temperature shows an general increasing trend from the beginning of January to mid-January.


## Problem 2
```{r}
library(tidyverse)
df <- tibble(
  random_samp = rnorm(10),
  is_positive = random_samp>0,
  vec_char = c("a", "b", "c", "a","b", "c","a","b", "c","a"),
  fac_vec = factor(c("male","other","female","female","male","other","male","female","male","female"))
)
df
```
 
```{r}
mean(df |> pull(random_samp))  
mean(df |> pull(is_positive))  
mean(df |> pull(vec_char))  
mean(df |> pull(fac_vec)) 
```

The mean function worked for numeric and logical vector but not for character and factor vector. 

```{r, results='hide', message=FALSE,warning=FALSE}
as.numeric(df |> pull(is_positive))
as.numeric(df |> pull(vec_char))
as.numeric(df |> pull(fac_vec))
```
The logical vector are all converted to 1 if it's True and 0 if it's false. It's working as it's computing the proportion of TRUE values.  

A error message appeared when trying to convert character to numeric data, "NA introduced by coercion". R doesn't know how to turn 'a','b','c' into numbers. That is why it failed when we try to take the mean.  

The factor vector are all converted to numeric vector, but the numbers are underlying integer codes (e.g., `"female" = 1`, `"male" = 2`, `"other" = 3` depending on alphabetical order).  
These codes are arbitrary labels, so their mean has no meaningful interpretation. That is why it failed when we try to take the mean.  

